Mapping factors associated with adolescents' emotion dysregulation: A text-mining based systematic review

### 2020-12-16 manuscript.Rmd
Finalize manuscript

abstract: |
  <!--RK: Wat is je onderzoeksvraag?-->Adolesence is a developmentally sensititve period for emotion dysregulation.
  Although substantial empirical research has addressed relevant risk factors, an integrating framework is lacking. This systematic review of 6305 papers uses text mining to map correlates of adolescents’ (ages 10-25) emotion dysregulation. One analysis focused on author-provided keywords; a second analysis parsed the abstracts. Terms were coded as putative “Causes”, “Outcomes”, “Protective factors”, or “Indicators”. We reported term/document frequency, and force-directed graphs of term co-occurrence. These results reﬂected constructs commonly featured in theory and narrative reviews, but also identified underrepresented terms. This research constitutes a first step towards integrating the literature, and illustrates how text mining reviews may complement narrative reviews. <!--may inspire future empirical and theoretical work. SH: I would end with a more memorable take-home about how text-mining can be useful to researchers (e.g., when topics are of multi-disciplinary interest but each field has their own preferred methods and/or terminology)-->

keywords          : "emotion regulation, adolescence, systematic review, text mining, machine learning"
wordcount         : "5400"
bibliography      : ["veni_sysrev.bib"]
floatsintext      : no
figurelist        : no
tablelist         : no
footnotelist      : no
linenumbers       : no
mask              : no
draft             : no
documentclass     : "apa6"
classoption       : "man,noextraspace"
output            : papaja::apa6_word
knit              : worcs:::cite_essential
---

```{r setup, include = FALSE}
library("papaja")
library(bibliometrix)
library(revtools)
library(bib2df)
library(knitr)
library(kableExtra)
library(cowplot)
run_everything = FALSE
```
<!-- Reviewer suggestions:
Kristin Buss -->

@@

<!-- Emotion regulation in adolescence is important -->
A key developmental challenge in adolescence is acquiring mature emotion
regulation skills [@croneUnderstandingAdolescencePeriod2012;
@zimmermannEmotionRegulationEarly2014]. Emotion regulation is integral to
mental health [@aldaoEmotionregulationStrategiesPsychopathology2010;
@schaferEmotionRegulationStrategies2017; @braetEmotionRegulationChildren2014]
and social functioning [@reindlSocializationEmotionRegulation2016]. For as many
as one in five, adolescence marks the onset of emotion
regulation-related mental illness, which can persist throughout the life course
[see @leeAdolescentMentalHealth2014]. 
It is therefore crucial to identify which
risk factors and environmental hazards render some adolescents more susceptible
to emotional difficulties than others. <!--SH: Main goal of paper -->
The present review set out to address this question.
Although there is an abundance of empirical work on adolescents' emotion regulation, the literature is somewhat fragmented, because the topic has been investigated from different (sub)disciplines.
As a result, there is a lack of unifying theory, and narrative reviews fall short of comprehensive coverage of the vast but diffuse literature [@bussTheoriesEmotionalDevelopment2019].

The present systematic review seeks to overcome this limitation by adopting a novel method -
text mining based bibliographic analysis - to map all factors currently
considered to be relevant to adolescent emotion regulation.
Traditional narrative literature reviews have been shown to be marred by reliance on (small) convenience samples of the literature, confirmation bias,
and an undue emphasis on positive results [@littellEvidencebasedBiasedQuality2008]. 
Text mining, by contrast, is uniquely suited to provide a comprehensive overview of the literature.
It can cover vastly greater corpora than a human reader,
and the process by which text mining gleans insights from the literature is more objective, transparant, and reproducible.
Where human readers might be inclined to structure their reading of the literature around established ideas,
the inherent "fairness" of text mining gives emerging themes in the literature a chance to come to the fore.
Text mining identifies important topics in a body of literature and reveals which constructs are often studied together.
This potentially reveals topics and connections missed by narrative reviewers,
thereby laying the groundwork for future empirical and theoretical work. <!-- CJ To fill out gaps, following up on promising constructs and unify heretofore disparate regions of the lit -->
<!-- SH: this implies that the main contribution of the paper is predominantly methodological is that the case? Is there some primary thesis/argument that will be advanced? It is really important to take a top-down approach to that messaging, beginning here in the first paragraph-->
<!-- Ergens anders: Similarly, although
there are several *relevant* theories, there is no specific theoretical
framework to guide research on adolescents' emotion regulation. Although theory
development is beyond the scope of this paper, we set out to generate a
proto-theoretical nomological network of the potential risk factors,
manifestations, and outcomes of emotion regulation in adolescence. We discuss
how this nomological network relates to existing relevant theory, and conclude
with recommendations for future theory-generating efforts.-->

@@

Adolescence is defined as a life stage ranging from pubertal onset to
adult-like independence [@steinbergAgeOpportunityLessons2014].
It is thus demarcated by both biological and socio-cultural factors.
It has been argued that,
due to accellerated pubertal onset and delayed transitions to adult roles,
adolescence now ranges from 10-24 years in the modern Western world
[@sawyerAgeAdolescence2018].

<!-- From Barriola:
* adolescence, more than adulthood or childhood, is characterized by frequent
and intense emotions (Larson et al. 1980; Larson and Lampmanpetraitis 1989).
* signi?cant psychosocial changes that call for an increase in sophistication
of ER strategies (Steinberg 2005)
* novel stressors associated with social relationships, including pressures to
conform to adult expectations relating to maturity, educational achievement,
gender appropriate conduct, and other socially competent behaviors (Yap et al.
2007)
* Additionally, adolescents have an increased orientation toward autonomy
(Spear 2000) including autonomous emotional regulation-->

Adolescence is increasingly seen as a developmentally sensitive period for
mature emotion regulation skills. During this period, children experience rapid
biological, cognitive, and social changes that prompt new emotional
experiences, and tax existing regulatory abilities
[@steinbergAgeOpportunityLessons2014]. At the same time, the staggered
development of motivational-emotional and regulatory brain circuits gives rise
to a maturity gap, where youngsters pursue new experiences in life and love,
without being fully prepared to cope with the emotional outcomes
[@croneUnderstandingAdolescencePeriod2012].
Due to these coalescing changes,
adolescents display a restricted repertoire of emotion regulation strategies
compared to younger or older individuals
[@zimmermannEmotionRegulationEarly2014]. Consequently, adolescents experience
more frequent and intense (negative) emotions than children or adults [see
@silkAdolescentsEmotionRegulation2003]. The literature thus paints a picture of
adolescence as a chrysalis for emotional development: Children enter this stage
with emotion regulation skills adapted to the challenges of childhood. During
adolescence, emotional systems are rearranged substantially. These changes
render adolescents temporarily vulnerable to emotional dysregulation, but
ultimately serve a functional role in acquiring mature emotion regulation
skills. Eventually, most adolescents emerge ready to take on the challenges of
early adulthood. Unfortunately, a sizeable proportion of adolescents instead
manifests enduring emotional difficulties [@leeAdolescentMentalHealth2014].

@@

The results of our text mining analysis of the empirical literature on adolescents' emotion regulation will be best understood against the backdrop of the existing theoretical landscape.
We distinguish the topics covered in this theoretical review visually to highlight new issues
in the subsequent text mining analysis.
This allows us to see in what respects the text mining analysis reflects existing theory,
and in what respects it complements it.
We provide only a summary theoretical review here, as other recent publications offer more detailed reviews of theories of emotional development [@bussTheoriesEmotionalDevelopment2019] and emotion regulation in adolescence [@riedigerEmotionRegulationAdolescence2014].
These prior reviews remark that,
although many theories are *relevant* for emotion regulation development in adolescence,
few have addressed this topic directly.
Fewer theories still have comprehensively addressed important predictors of development in this life stage,
and none directly guide contemporary research in the field [see @bussTheoriesEmotionalDevelopment2019].
There is thus a need for more integrative theory formation [see @riedigerEmotionRegulationAdolescence2014].
Inductive analysis of the empirical literature may constitute a first step to this end.

Among the oldest relevant theories of adolescents' emotional development is Hall's notion of "storm and stress" [see
@arnettAdolescentStormStress1999]. It describes how hormonal changes diminish
self-control and increase reactivity, leading to difficulties in emotion
regulation, conflict with parents, and risky behavior.
In modern theories, this notion of diminished self-control and increased emotional reactivity has been retained - although it is now considered to be a normative change that facilitates emotional maturation
[@croneUnderstandingAdolescencePeriod2012], at the risk of emotional
disturbance [@arnettAdolescentStormStress1999; @leeAdolescentMentalHealth2014].

@@

One example of a normative theory of emotional development in early childhood is presented by Sroufe [-@sroufeEmotionalDevelopmentOrganization1995].
Sroufe argues that, as children grow older,
their increasing self-regulatory abilities
drive a transition from external emotion regulation by primary caregivers
towards autonomous emotion regulation. This theory focuses on two drivers of
development: social and cognitive influences. Social influences mainly occur
through parental co-regulation, parenting behaviors, and parent-child
attachment. Cognitive influences occur through the development of the central
nervous system (CNS), cognition, and self-regulation.

This emphasis on social influences and neurocognitive development is reflected in many theories.
Some of these focus primarily on neurocognitive development.
For example, polyvagal theory closely links emotional experience
- and regulation - to autonomous nervous system functioning
[@porgesOrientingDefensiveWorld1995].
A more recent neurocognitive theory relevant to adolescents' emotion regulation development is Crone
and Dahl's model of social-affective engagement and goal flexibility
[@croneUnderstandingAdolescencePeriod2012]. Like other authors 
[e.g., @caseyBrakingAcceleratingAdolescent2011; @craccoEmotionRegulationChildhood2017],
Crone and Dahl argue that the developmental
asymmetry between motivational and inhibitory brain circuits gives rise to a
"maturity gap" in middle adolescence. 
What distinguishes Crone and Dahl from related writings
is the greater attention to adolescents' diverging destinies:
the question why some youngsters flourish while others languish.
They argue that adolescents' cognitive engagement is dynamically responsive to
social and motivational goal salience. This flexibility, on the one hand,
prepares adolescents to effectively engage cognitive systems in novel
challenging situations in a way that facilitates developing mature regulatory
abilities. On the other hand, it places them at risk to act impulsively in
pursuit of peer approval. In terms of explanans,
this theory focuses primarily on cognitive factors
and the role of peers, with less attention to factors such as parenting.
Although this theory focuses on adolescence specifically,
an important limitation is that it only tangentially relates to emotion regulation.

Other theories emphasize socialization to a greater extent.
One such theory is Morris' tripartite model, which focuses specifically on parents' role in emotion regulation development [-@morrisRoleFamilyContext2007].
It describes three pathways through which parents shape emotion regulation
development: Through observation and modeling, parenting practices, and the
emotional family climate, which in turn involves attachment and marital
relationship quality. Morris and colleagues also emphasize the relevance of
fathers and siblings, and others have adapted the tripartite model to describe
the role of peers in adolescents' emotion regulation development
[@reindlSocializationEmotionRegulation2016].
A more abstract take on socialization is found in @holodynskiDevelopmentEmotionsEmotion2006's
internalization model of emotional development. 
Unique aspects emphasized in this theory are the interplay between emotion and communication,
and the internalization of the cultural symbolic function of emotion.
These theories explain socialization in great detail,
but one potential limitation is that tey are not embedded in a larger unifying framework.
<!--Like Sroufe, Holodynski describes an age graded transition from
interpersonal to intrapersonal emotion regulation.
In doing so, Holodynski applies Vygotsky's theory of development to the domain of
emotion, and presents a detailed integrated model of emotional experience and
regulation.-->

Looking beyond the developmental literature,
theories of the phenomenon of emotion regulation offer additional insight into
intra-individual drivers of emotion regulation development.
Perhaps the most influential theory is Gross'
[-@grossHandbookEmotionRegulation2013] process model, which describes the
process of emotion regulation from eliciting cue to ultimate response.
According to Gross, individuals use strategies to modulate the different stages of this process,
consciously or otherwise. Culture plays a role too, as the effectiveness, desireability, and consequences of different strategies depend on the social context [see
@bariolaChildAdolescentEmotion2011]. Similar to Gross' theory, the social
information processing theory of emotion also describe the role of cognitive
processes and strategies in emotion experience and regulation
[@lemeriseIntegratedModelEmotion2000].

Finally, given our interest in risk factors that render adolescents susceptible
to emotional difficulties, we should mention two theories widely invoked to frame research on developmental influences: Bronfenbrenner's bioecological model, and Sameroff's
transactional model. Bronfenbrenner's model
[@bronfenbrennerBioecologicalModelHuman2007] describes how the environment
shapes individual development. Each individual is imbued with certain
biological predispositions, and develops over time, in interaction with
contextual influences. These influences range from the microsystem, composed of
those people close to the individual, to the macrosystem, consisting of
indirect political and economic influences, to the exosystem, consisting of
cultural norms and values. Sameroff's model is somewhat compatible with the
bioecological model [@sameroffUnifiedTheoryDevelopment2010], but focuses more
on the interplay between nature and nurture. It conceptualizes development as a
product of reciprocal influences between the child and the environment.
Sameroff distinguishes between proximal influences, for instance, through
parents (the microsystem in Bronfenbrenner's model), and distal influences;
structural factors indirectly shaping development, like socio-economic status,
schools, and the community (macro- and exosystem). With increasing age, distal
influences gradually gain ground on proximal influences.
These theories are broad enough to contextualize any developmental study,
but limited by a lack of specificity, which hampers their utility in generating hypotheses.
<!-- RK: What do you take from these broad perspectives for your RQ?-->
<!--SH I am not sure what contribution this paragraph ultimately makes - you should consider condensing it substantially to just note that these theories have something to say about emotional development but are not adolescent-specific-->
<!--Theories like this are particularly relevant to a fine-grained
understanding of the process of emotion regulation, but their developmental
relevance has unfortunately not yet been considered (REF Buss). -->
<!--Parents are widely considered to be the primary proximal influence driving
emotion regulation development.
Yet several limitations remain.
@coe-odessEmergentEmotionsAdolescence2019;
instead takes an inductive approach to map the correlates of adolescents'
emotion regulation based on all available literature.

@@

@@

<!--Consider splitting this up and putting the shortcomings next to the existing theories-->
Despite the abundance of relevant theory,
the existing literature has several limitations when the goals is to
identify which
risk factors render some adolescents more susceptible
to maladaptive emotional development than others.
One crucial shortcoming is that most theories lack of explicit attention to the life stage of adolescence.
It is questionable whether theories focused on different ages can be generalized to adolescents,
because this stage differs qualitatively from infancy, childhood, and adulthood [@bariolaChildAdolescentEmotion2011].
A related limitation is that many relevant theories lack
an explicit developmental component <!--, and thus lack a well-substantiated
understanding of developmental processes--> [see
@bussTheoriesEmotionalDevelopment2019;
@croneUnderstandingAdolescencePeriod2012]. 
Perhaps most importantly, however, these theories
vary widely in scope: Some are broad but somewhat non-specific, such as the biopsychosocial models of Bronfenbrenner and Sameroff, and others describe a specific phenomenon, but lack a broader perspective.
It would be beneficial to bridge these levels of analysis.
Based on these shortcomings,
there appears to be a need for more theory generation,
in order to provide a unified framework that could guide future empirical work.
<!--Such a theory should focus on -->
The present study lays the groundwork for such theory development,
by providing an inductive conceptual
review of the field.
Concepts covered in the preceding theoretical review will be tagged in the text mining analysis,
to highlight the contribution of the inductive approach.

@@

## Bottom-up insights from narrative reviews

Whereas theory provides a top-down understanding of an area of research,
literature reviews can provide a bottom-up, inductive understanding. Given the
noted absence of a single overarching theory, reviews might provide additional
insight into factors considered relevant in the etiology of adolescents'
emotion regulation competencies. For example, a review by
@bariolaChildAdolescentEmotion2011 addresses individual psychological factors
such as temperament, and biological factors like neurocognitive development and
genetic predispositions. Furthermore, the authors address proximal social influences,
including parents, teachers, and peers, and specific mechanisms through which
these individuals exert their influence, including parenting and modeling.
Finally, they address distal factors, including the role of culture and the
media. Barriola and colleagues also 
<!--most notably, they call for a robust theoretical framework to guide future
research.--> call for future research to address parents' role beyond early
childhood.
They further argued that the role of fathers in emotion socialization warrants further scrutiny - a topic that is now receiving more attention [@vanlissaRoleFathersMothers2019;
@vanlissaMothersFathersQuantitative2020].

>>>>

abstract: |
  Adolesence is a developmentally sensititve period for emotion dysregulation.
  Although substantial empirical research has addressed relevant risk factors,
  the literature is fragmented across subdisciplines, and
  an integrating framework is lacking.
  This systematic review of 6305 papers used text mining to map correlates of adolescents’ (ages 10-25) emotion dysregulation. First, we established a baseline of relevant terms gleaned from theory and recent narrative reviews. Then, we conducted two text mining analyses to examine term/document frequency and co-occurrence in author-provided keywords and abstracts. Results reﬂected constructs commonly featured in theory and narrative reviews, but also identified underrepresented themes. This research constitutes a first step towards integrating the literature, and illustrates how text mining reviews may complement narrative reviews. <!--may inspire future empirical and theoretical work. SH: I would end with a more memorable take-home about how text-mining can be useful to researchers (e.g., when topics are of multi-disciplinary interest but each field has their own preferred methods and/or terminology)-->

wordcount         : "5400"
bibliography      : ["veni_sysrev.bib"]
floatsintext      : no
figurelist        : no
tablelist         : no
footnotelist      : no
linenumbers       : no
mask              : no
draft             : no
documentclass     : "apa6"
classoption       : "man,noextraspace"
output            : papaja::apa6_pdf
knit              : worcs:::cite_essential
---

```{r setup, include = FALSE}
library("papaja")
library(bibliometrix)
library(revtools)
library(bib2df)
library(knitr)
library(kableExtra)
library(cowplot)
run_everything = FALSE
notingraph <- list()
```
<!-- Reviewer suggestions:
Kristin Buss -->

@@

<!-- Emotion regulation in adolescence is important -->
A key developmental challenge in adolescence is acquiring mature emotion
regulation skills [@croneUnderstandingAdolescencePeriod2012;
@zimmermannEmotionRegulationEarly2014]. Emotion regulation is integral to
mental health [@aldaoEmotionregulationStrategiesPsychopathology2010;
@schaferEmotionRegulationStrategies2017; @braetEmotionRegulationChildren2014]
and social functioning [@reindlSocializationEmotionRegulation2016]. For as many
as one in five individuals, adolescence marks the onset of emotion
regulation-related mental illness, which can persist throughout the life course
[see @leeAdolescentMentalHealth2014]. 
It is therefore crucial to identify which
risk factors and environmental hazards render some adolescents more susceptible
to emotional difficulties than others. <!--SH: Main goal of paper -->
Although there is an abundance of empirical work on adolescents' emotion regulation, the literature is somewhat fragmented, because the topic has been investigated from different (sub)disciplines.
As a result, there is a lack of unifying theory,
and intrinsic limitations of narrative reviews prohibit comprehensive coverage of the vast but diffuse literature [@bussTheoriesEmotionalDevelopment2019].

The present systematic review seeks to overcome these limitations by adopting a novel method:
the text mining based systematic review.
We set out to map the constructs associated with adolescent emotion regulation based on a systematic review of the empirical literature.
Traditional narrative literature reviews have been shown to be marred by reliance on small convenience samples of the literature, confirmation bias,
and an undue emphasis on positive results [@littellEvidencebasedBiasedQuality2008]. 
Text mining, by contrast, is uniquely suited to provide a comprehensive overview of the literature.
It can cover vastly greater corpora than a human reader,
and the process by which text mining gleans insights from the literature is more objective, transparant, and reproducible.
Where human readers might be inclined to structure their reading of the literature around established ideas,
the inherent "fairness" of text mining gives emerging themes in the literature a chance to come to the fore.
Text mining identifies important topics in a body of literature and reveals which constructs are often studied together.
This potentially reveals topics and connections missed by narrative reviewers,
thereby laying the groundwork for future empirical and theoretical work. <!-- CJ To fill out gaps, following up on promising constructs and unify heretofore disparate regions of the lit -->
<!-- SH: this implies that the main contribution of the paper is predominantly methodological is that the case? Is there some primary thesis/argument that will be advanced? It is really important to take a top-down approach to that messaging, beginning here in the first paragraph-->
<!-- Ergens anders: Similarly, although
there are several *relevant* theories, there is no specific theoretical
framework to guide research on adolescents' emotion regulation. Although theory
development is beyond the scope of this paper, we set out to generate a
proto-theoretical nomological network of the potential risk factors,
manifestations, and outcomes of emotion regulation in adolescence. We discuss
how this nomological network relates to existing relevant theory, and conclude
with recommendations for future theory-generating efforts.-->

@@

Adolescence is defined as a life stage ranging from pubertal onset to
adult-like independence [@steinbergAgeOpportunityLessons2014].
It is thus demarcated by both biological and socio-cultural factors.
It has been argued that,
due to accellerated pubertal onset and delayed transitions to adult roles,
adolescence now ranges from 10-24 years in the modern Western world
[@sawyerAgeAdolescence2018].
<!-- From Barriola:
* adolescence, more than adulthood or childhood, is characterized by frequent
and intense emotions (Larson et al. 1980; Larson and Lampmanpetraitis 1989).
* signi?cant psychosocial changes that call for an increase in sophistication
of ER strategies (Steinberg 2005)
* novel stressors associated with social relationships, including pressures to
conform to adult expectations relating to maturity, educational achievement,
gender appropriate conduct, and other socially competent behaviors (Yap et al.
2007)
* Additionally, adolescents have an increased orientation toward autonomy
(Spear 2000) including autonomous emotional regulation-->
Adolescence is increasingly seen as a developmentally sensitive period for
mature emotion regulation skills. During this period, children experience rapid
biological, cognitive, and social changes that prompt new emotional
experiences, and tax existing regulatory abilities
[@steinbergAgeOpportunityLessons2014]. At the same time, the staggered
development of motivational-emotional and regulatory brain circuits gives rise
to a maturity gap [@croneUnderstandingAdolescencePeriod2012]. 
This leads youngsters to pursue new experiences in life and love,
without being fully prepared to cope with the emotional outcomes.
Due to these coalescing changes,
adolescents display a restricted repertoire of emotion regulation strategies
compared to younger or older individuals
[@zimmermannEmotionRegulationEarly2014]. Consequently, adolescents experience
more frequent and intense (negative) emotions than children or adults [see
@silkAdolescentsEmotionRegulation2003]. The literature thus paints a picture of
adolescence as a chrysalis for emotional development: Children enter this stage
with emotion regulation skills adapted to the challenges of childhood. During
adolescence, emotional systems are rearranged substantially. These changes
render adolescents temporarily vulnerable to emotional dysregulation, but
ultimately serve a functional role in acquiring mature emotion regulation
skills. Eventually, most adolescents emerge ready to take on the challenges of
early adulthood. Unfortunately, a sizeable proportion of adolescents instead
manifests enduring emotional difficulties [@leeAdolescentMentalHealth2014].

@@

The results of our text mining analysis of the empirical literature are best understood against a backdrop of the existing theoretical landscape.
We therefore provide a summary theoretical review.
Other recent publications offer more detailed reviews of theories of emotional development [@bussTheoriesEmotionalDevelopment2019] and emotion regulation in adolescence [@riedigerEmotionRegulationAdolescence2014].
The correlates of emotion regulation difficulties discussed in the theoretical review below
are visually summarized in the left panel of Figure \@ref(fig:baseline)^[Based on an author-curated list of keywords, caterogized using the dictionary method introduced in Analysis 1.].

```{r baseline, warning=FALSE, eval = TRUE, out.width = "100%", fig.cap="Correlates of emotion regulation difficulties according to theory (left panel) and narrative reviews (right panel). Transparent circles indicate constructs also represented in the theory."}
knitr::include_graphics("baseline_network.png") 
#fig_svg <- cowplot::ggdraw()+cowplot::draw_image("s1_varimp.svg")
#plot(fig_svg)
keyw <- function(x){
  cat(x)
}
```

Among the oldest relevant theories of adolescents' emotional development is Hall's notion of "storm and stress" [see
@arnettAdolescentStormStress1999]. It describes how `r keyw("hormonal")` changes diminish
`r keyw("self-control")` and increase `r keyw("reactivity")`, leading to difficulties in emotion
regulation, `r keyw("conflict with parents")`, and `r keyw("risky behavior")`.
In modern theories, this notion of diminished self-control and increased emotional reactivity has been retained - although it is now considered to be a normative change that facilitates emotional maturation
[@croneUnderstandingAdolescencePeriod2012], at the risk of emotional
disturbance [@arnettAdolescentStormStress1999; @leeAdolescentMentalHealth2014].

@@

One example of a normative theory of emotional development in early childhood is presented by Sroufe [-@sroufeEmotionalDevelopmentOrganization1995].
Sroufe argues that, as children grow older,
their increasing `r keyw("self-regulatory")` abilities
drive a transition from external emotion regulation by `r keyw("primary caregivers")`
towards autonomous emotion regulation. This theory focuses on two drivers of
development: `r keyw("social")` and `r keyw("cognitive influences")`. Social influences mainly occur
through `r keyw("parental co-regulation")`, `r keyw("parenting behaviors")`, and parent-child
`r keyw("attachment")`. Cognitive influences occur through the development of the `r keyw("central
nervous system")` (CNS), `r keyw("cognition")`, and `r keyw("self-regulation")`.

This emphasis on social influences and `r keyw("neurocognitive development")` is reflected in many theories.
Some of these focus primarily on neurocognitive development.
For example, `r keyw("polyvagal theory")` closely links `r keyw("emotional experience")`
- and regulation - to autonomous nervous system functioning
[@porgesOrientingDefensiveWorld1995].
A more recent neurocognitive theory relevant to adolescents' emotion regulation development is Crone
and Dahl's model of social-affective engagement and `r keyw("goal flexibility")`
[@croneUnderstandingAdolescencePeriod2012]. Like other authors 
[e.g., @caseyBrakingAcceleratingAdolescent2011; @craccoEmotionRegulationChildhood2017],
Crone and Dahl argue that the developmental
asymmetry between `r keyw("motivational")` and inhibitory `r keyw("brain")` circuits gives rise to a
"maturity gap" in middle adolescence. 
What distinguishes Crone and Dahl from related writings
is the greater attention to adolescents' diverging destinies:
the question why some youngsters flourish while others languish.
They argue that adolescents' cognitive engagement is dynamically responsive to
social and motivational `r keyw("goal salience")`. This flexibility, on the one hand,
prepares adolescents to effectively engage cognitive systems in novel
challenging situations in a way that facilitates developing mature regulatory
abilities. On the other hand, it places them at risk to act impulsively in
pursuit of `r keyw("peer approval")`. In terms of explanans,
this theory focuses primarily on cognitive factors
and the role of peers, with less attention to factors such as parenting.
Although this theory focuses on adolescence specifically,
an important limitation is that it only tangentially relates to emotion regulation.

Other theories emphasize socialization to a greater extent.
One such theory is Morris' tripartite model, which focuses specifically on `r keyw("parents")`' role in emotion regulation development [-@morrisRoleFamilyContext2007].
It describes three pathways through which parents shape emotion regulation
development: Through observation and `r keyw("modeling")`, `r keyw("parenting practices")`, and the
emotional `r keyw("family climate")`, which in turn involves attachment and `r keyw("marital
relationship")` quality. Morris and colleagues emphasize the relevance of `r keyw("mothers")`, 
`r keyw("fathers")`, and siblings.
Others have adapted the tripartite model to describe
the role of peers in adolescents' emotion regulation development
[@reindlSocializationEmotionRegulation2016].
A more abstract take on socialization is found in @holodynskiDevelopmentEmotionsEmotion2006's
internalization model of emotional development. 
Unique aspects emphasized in this theory are the interplay between emotion and `r keyw("communication")`,
and the internalization of the `r keyw("cultural")` symbolic function of emotion.
These theories explain the process of socialization in great detail,
but one potential limitation is that they are not embedded in a larger unifying framework.
<!--Like Sroufe, Holodynski describes an age graded transition from
interpersonal to intrapersonal emotion regulation.
In doing so, Holodynski applies Vygotsky's theory of development to the domain of
emotion, and presents a detailed integrated model of emotional experience and
regulation.-->

Looking beyond the developmental literature,
theories of the phenomenon of emotion regulation offer additional insight into
intra-individual drivers of emotion regulation development.
Perhaps the most influential theory is Gross'
[-@grossHandbookEmotionRegulation2013] process model, which describes the
process of emotion regulation from eliciting cue to ultimate response.
According to Gross, individuals use `r keyw("strategies")` to modulate the different stages of this process,
consciously or otherwise. `r keyw("Culture")` plays a role too, as the effectiveness, desireability, and consequences of different strategies depend on the social context [see
@bariolaChildAdolescentEmotion2011]. Similar to Gross' theory, the social
information processing theory of emotion also describe the role of cognitive
processes and strategies in emotion experience and regulation
[@lemeriseIntegratedModelEmotion2000].

Finally, given our interest in risk factors that render adolescents susceptible
to emotional difficulties, we should mention two theories widely invoked to frame research on developmental influences: Bronfenbrenner's bioecological model, and Sameroff's
transactional model.
Bronfenbrenner's model
[@bronfenbrennerBioecologicalModelHuman2007] describes how the `r keyw("environment")`
shapes individual development. Each individual is imbued with certain
`r keyw("biological")` predispositions, and develops over time, in interaction with
contextual influences. These influences range from the microsystem, composed of
those people close to the individual, to the macrosystem, consisting of
indirect `r keyw("political")` and `r keyw("economic")` influences, to the exosystem, consisting of
cultural norms and values. Sameroff's model is somewhat compatible with the
bioecological model [@sameroffUnifiedTheoryDevelopment2010], but focuses more
on the interplay between nature and nurture. It conceptualizes development as a
product of reciprocal influences between the child and the environment.
Sameroff distinguishes between proximal influences, for instance, through
parents (the microsystem in Bronfenbrenner's model), and distal influences;
structural factors indirectly shaping development, like socio-economic status,
`r keyw("schools")`, and the `r keyw("community")` (macro- and exosystem). With increasing age, distal
influences gradually gain ground on proximal influences.
These theories are broad enough to contextualize any developmental study,
but lack specificity, which curtails their utility in generating hypotheses.
<!-- RK: What do you take from these broad perspectives for your RQ?-->
<!--SH I am not sure what contribution this paragraph ultimately makes - you should consider condensing it substantially to just note that these theories have something to say about emotional development but are not adolescent-specific-->
<!--Theories like this are particularly relevant to a fine-grained
understanding of the process of emotion regulation, but their developmental
relevance has unfortunately not yet been considered (REF Buss). -->
<!--Parents are widely considered to be the primary proximal influence driving
emotion regulation development.
Yet several limitations remain.
@coe-odessEmergentEmotionsAdolescence2019;
instead takes an inductive approach to map the correlates of adolescents'
emotion regulation based on all available literature.

@@

<!--
Note that the left panel of Figure \@ref(fig:baseline) visually summarizes the correlates of emotion regulation difficulties (labeled: *Dysregulation*) according to our theoretical review.
The left panel of Figure \@ref(fig:baseline) summarizes key constructs covered by the theories reviewed above. 
In the subsequent text mining analysis, we will distinguish the topics covered in this theoretical review visually to highlight unique contributions of our approach.
This allows us to see in what respects the text mining analysis reflects existing theory,
and in what respects it complements it.-->

@@

Despite the abundance of theory *relevant* to emotion regulation in adolescence,
the theoretical literature has several limitations.
First, few theories have explicitly addressed the life stage of adolescence. This life stage differs qualitatively from infancy, childhood, and adulthood [@bariolaChildAdolescentEmotion2011].
It is therefore questionable whether more general theories, or those focused on different age groups, can be generalized to adolescents.
<!--Relatedly, many relevant theories lack an explicit developmental component, which renders them and thus lack a well-substantiated
understanding of developmental processes[see
@bussTheoriesEmotionalDevelopment2019;
@croneUnderstandingAdolescencePeriod2012]. -->
Furthermore, few theories have comprehensively addressed important predictors of development in this life stage,
and none directly guide contemporary research in the field [see @bussTheoriesEmotionalDevelopment2019; @riedigerEmotionRegulationAdolescence2014].
Finally, available theories vary widely in scope:
Some are broad but somewhat non-specific, and others describe a specific phenomenon in detail,
but lack a broader perspective.
It would be beneficial to bridge these levels of analysis.
In sum, there is a need for more integrative theory formation,
in order to provide a unified framework that could guide future empirical work.<!--[see @riedigerEmotionRegulationAdolescence2014].
Inductive analysis of the empirical literature may constitute a first step to this end.-->
<!--Such a theory should focus on -->
The present study lays the groundwork for such theory development,
by providing a text-mining based inductive systematic review of the field.

@@

## Prior narrative reviews

Whereas theory provides a top-down frame of reference for understanding a phenomenon,
literature reviews can provide a bottom-up, inductive understanding.
Given the noted absence of a single overarching theory, reviews might provide additional
insight into factors considered relevant in the etiology of adolescents'
emotion regulation competencies. 
Innovative text mining methods are promising in this respect, because they can comprehensively review the available literature.
Narrative reviews, however, constitute the status quo.
We will therefore examine how two excellent narrative reviews on emotion regulation in adolescence complement the aforementioned theoretical literature [@coe-odessEmergentEmotionsAdolescence2019; @bariolaChildAdolescentEmotion2011].
The right panel of Figure \@ref(fig:baseline) visually distinguishes constructs  uniquely covered by these narrative reviews,
relative to the issues also addressed in the theoretical literature (left panel).

<!--
Note that the left panel of Figure \@ref(fig:baseline) visually summarizes the correlates of emotion regulation difficulties (labeled: *Dysregulation*) according to our theoretical review.
The left panel of Figure \@ref(fig:baseline) summarizes key constructs covered by the theories reviewed above. 
In the subsequent text mining analysis, we will distinguish the topics covered in this theoretical review visually to highlight unique contributions of our approach.
This allows us to see in what respects the text mining analysis reflects existing theory,
and in what respects it complements it.-->

The seminal review by
@bariolaChildAdolescentEmotion2011 addresses individual psychological factors
such as temperament, and biological factors like neurocognitive development and
genetic predispositions. Furthermore, the authors address proximal social influences,
including parents, teachers, and peers, and specific mechanisms through which
these individuals exert their influence, including parenting and modeling.
Finally, they address distal factors, including the role of culture and the
media. Barriola and colleagues also 
<!--most notably, they call for a robust theoretical framework to guide future
research.--> call for future research to address parents' role beyond early
childhood.
They further argued that the role of fathers in emotion socialization warrants further scrutiny - a topic that is now receiving more attention [e.g., see @vanlissaRoleFathersMothers2019;
@vanlissaMothersFathersQuantitative2020].

A more recent review by @coe-odessEmergentEmotionsAdolescence2019 offers a nuanced
discussion of several issues that complement prior publications in this field.
For instance, the authors cover the implications of physiological changes in
detail, including neurocognitive development and pubertal maturation. As
adolescents become increasingly individuated, conflict with parents peaks. This
conflict, in turn, impacts emotion regulation, both in terms of day-to-day mood
variability and dispositional emotion dysregulation [see also
@vanlissaCostEmpathyParentadolescent2017].
Compared to children, adolescents also become increasingly oriented towards peers.
This includes increased
sensitivity to social status and norms, along with the associated increase in
peer pressure and risk taking. Pubertal development also initiates the
emergence of sexual and romantic behavior, and the intensification of both
biological sex differences and gender stereotyped behavior. With regard to the
process of emotion regulation, Coe-Odess and colleagues discuss the importance
of strategies, and point out that, in addition to negative emotionality,
positive emotionality also peaks in adolescence. Relatedly, the authors
describe how hormonal changes precipitate an increased stress response, which
might help explain adolescents' greater susceptibility to emotion
dysregulation. Finally, going beyond the implications of cognitive development
discussed in other publications, the authors discuss how an greater capacity
for abstract thought relates to identity formation - a key challenge in
adolescence [@meeusStudyAdolescentIdentity2011] - and to increased emotional
understanding, and by extension, empathy [see also
@vanlissaLongitudinalInterplayAffective2014].
